%   Filename    : acknowledgment.tex 
\begin{center}
	\textbf{Acknowledgment}
\end{center}

\noindent\textbf{I. General Acknowledgment} \\
\vspace{-0.5cm}

The researchers would like to extend their heartfelt gratitude to the significant people who were part of this journey. These people extended their expertise, support, and guidance, which made this whole process bearable and fulfilling.

First and foremost, to the Almighty God who gives us the strength to never lose hope throughout the challenging moments in our journey; for the wisdom and perseverance He bestowed upon us to accomplish this paper. All for His unconditional love, mercy, and grace. All glory and praise to Him.

To our research adviser who provided guidance and support to the researchers, and for the help when things were hard. Dr. Francis Dimzon – thank you for guiding us throughout this journey. We would like to thank you for assuring us, especially when we were about to turn around halfway. We would like to extend our greatest gratitude for the consistent supervision by providing us with valuable feedback and insightful comments that aided us in our methodology and in completing this paper.

To our co-adviser who was willing to tour us around the hatchery facility and helped us come up with this study. Thank you, Dr. Victor Marco Emmanuel Ferriols, for sharing your knowledge and wisdom with us researchers. Thank you for answering our questions and extending your time with us. Additionally, the most important part of the research is the samples — the blood cockle samples — thank you for providing them willingly. You served as an inspiration for us to apply solutions to real-world problems, especially in aquaculture.

To the research associates and hatchery staff of the UPV Hatchery, Ms. Allena Arteta, Ms. LC May Gasit, Ms. Shiela Untalan, and the hatchery staff, Mr. Paul Andre Lopez and Mr. Joel M. Fabrigas, for assisting us in handling the blood cockles, teaching us how to identify their sex, spawning, and dissection. A big thanks to Ms. Allena Arteta for letting us borrow the camera stand that we used in the entire data gathering process. Your constant warmth and assistance went beyond learning the basics; instead, we learned a whole lot more, especially the importance of creating solutions for aquaculture practices.

Lastly, to everyone who directly or indirectly contributed to this thesis, whether through words of encouragement, technical advice, or simply by believing in us, we offer our sincerest gratitude. This study would also not have been possible without you.

\vspace{0.5cm}
\noindent\textbf{i. Adricula's Acknowledgment} \\
\vspace{-0.5cm}

I would like to express my deepest gratitude to my beloved family who supported me throughout my academic journey. My Mama Madelene and Papa Juvy, who have made countless sacrifices to make me the person I am today. To my sister, Aianna, and cousins Jayson, Joana, and Joel, who have helped me unwind after long and stressful days of academic load. To my aunt, Tita Lybel – thank you for believing in me since day one. You are always there to guide me in every decision I make. And to my Lola Lydia, who has been my inspiration in finishing my studies. Your reminders to eat more and stay safe have always kept me grounded and cared for, and I appreciate the extra allowance you always give me whenever I go back to Miagao. Without such a supportive family behind me, I couldn’t have done this without you.

To my pretty college friends, Arianne, Gliezel, Karielle, Kane, Kzlyr, and Sharah, thank you for making my college experience bearable. The laughter and joy that we shared together lessened the stress I felt. Here’s to all the rants, tsismis, and bangs memories we had — look how far we’ve come. I will forever cherish the memories we made together through ups and downs. Thank you for the encouragement along the way and just for being there by my side.

To my co-researchers, Gliezel and Kane, this would not have been possible without you. Your insights and passion for this research inspired me along the way. Thank you for the teamwork, initiatives, and hard work. I will miss the dorm to hatchery to computer lab to dorm adventures with you!

To my second home, UPV Balay Gumamela, our dorm manager, and all the staff — thank you for checking up on me, always making me feel welcome, and ensuring that we had a comfortable stay.

And lastly, to the people who believed in me and have always been there for me every step of the way, even when life was harsh — I am forever grateful for your unwavering encouragement, support, and just for believing that I can make things possible.

\vspace{0.5cm}
\noindent\textbf{ii. Pajarilla's Acknowledgment} \\
\vspace{-0.5cm}

To my mother in heaven, Mama Eva, I made it through the last chapter of this paper. I felt your boundless love and support every step of the way. You will always be my source of strength and inspiration, even when hope seemed bleak and weary. The different storms that were pushed my way — I stood because I know you believed in me and wanted the best for me and my three siblings.

Greatest love and appreciation to my siblings – my source of boundless joy. Your smiles, laughter, and warmth gave me a boost of energy that adenosine triphosphate can’t surpass. Gil Ivan, Gillian Gella, and my small baby brother Gian Godric – no words can express how thankful I am for your existence and love, as it sparked perseverance and grit in my heart.
To the people who cared and looked after me, thank you, Tito Rem and Auntie Maricel, for the financial support and encouragement. Thank you for believing in me and the things that I can achieve.

My heart is full of gratitude and appreciation to the Gumamela Girls Dormers. I cherish all the laughter, joy, sadness, ennui, anger, fear, embarrassment, anxiety, envy, and disgust. Kidding aside, I am forever thankful for the bond and friendship we made, the spontaneous moments, and baybay sessions, admiring the beauty of nature. Arianne, Briana, Kane, Karielle, Kzlyr, Sharah – thank you for the words of encouragement, for listening, and for your companionship.

I would like to extend my greatest appreciation to my co-researchers, Kane and Briana, who trailed the 100 steps from and to the hatchery facility. I appreciate the moments and laughter that we spent together, from ideation, spending almost the whole day in the computer laboratory to train our model, to crafting this final paper. We went through doubts and a series of burnout moments, but your presence made everything lighter.

My love and appreciation are offered to my high school and Senior High School friends who listened to my “I can’t do this anymore” moments. Thank you, Jara, Francine, Nellen, Mary, and Nyle, for cheering me on and believing in my potential, even when things sometimes got so dim. Thank you for reminding me that there is always light at the end of the tunnel. You all are my support system, cheering squad, and parents I met along the way.

My gratitude extends to the people inside the university who served as family. To Ma’am Tess Geonanga, Ma’am Paula Khryss Ushiyama, Balay Gumamela dormer manager and staff, and the UPV Virgils — thank you for the support, for believing in me, and for the nutritional food that helped me brainstorm, conduct, and write this study.
To everyone who supported me, thank you so much, and words can’t express how grateful I am.

\vspace{0.5cm}
\noindent\textbf{iii. Vito's Acknowledgment} \\
\vspace{-0.5cm}

To my beloved family – Mama, Papa, Mommy, Lola Jo, Uncle Mike, Tita Emie, Lola Shirley, and Lola Terry – thank you for your unconditional love, patience, and never-ending encouragement. Your strength and support have been the backbone of everything I’ve achieved. Your belief in me has always kept me grounded and motivated, even when I doubted myself. Thank you for the constant prayers, thoughtful check-ins, and warm words that lifted me during the hardest times. Your financial support has also played an instrumental role in helping me pursue this journey, and I am deeply grateful.

To my lovely (but sometimes annoying) siblings – Hariette, Vladimir, Michaela, Regina, and Daniel – words can’t fully express how much you mean to me. Thank you for being my constant source of joy, laughter, and sanity. Through all the sibling chaos, inside jokes, and heart-to-heart conversations, you've kept me connected to the things that truly matter. Whether it was a random meme shared to make me laugh, a quick call just to check in, or simply being there in quiet support, you all reminded me that no matter where I was, I was never truly alone.

To my ever-reliable and ever-beautiful Gumamela Girls – Briana, Gliezel, Arianne, Sharah, Kai, and Kzlyr – thank you for the love, the chika, and the comfort. You all made even the most stressful days feel manageable. You’ve truly been my home away from home. Special mention to Briana and Gliezel, my fellow researchers, who stood by me not only as friends but as teammates navigating this academic journey together. Your dedication, the leg-torturing stairs we climbed to the hatchery for data gathering, and late-night grind sessions will always be one of my favorite parts of this experience.

To the UPV Virgils, thank you for being the silent heroes behind my survival. Your monthly grocery supplies kept me fed, sane, and functioning. I probably owe half of this thesis to the snacks.

To the Balay Gumamela dorm staff, thank you for being my guardians during my stay at the dorm. Your support, care, and constant presence made my time away from home more comfortable and secure. I will always be grateful for the way you looked out for me and made me feel safe.

To my very cutesy, very demure, very mindful pet dogs, Chewy and Max – thank you for the joy and companionship you’ve brought into my life. Max, though you are no longer with us, your memory continues to live on in my heart. Chewy, thank you for being my ever-present companion, sharing both the quiet, loud, and the playful moments.

To chonky cat Mother Litob, your cuteness and presence in the hatchery during our data gathering made the 100-step stairs bearable. My tiredness would instantly vanish whenever I petted you.

This journey has been a collective effort – held up by the love, wisdom, laughter, and resilience of those around me. I am endlessly grateful for each of you.



